---
title: "Using linregpackage"
author: "Venkata Balaji Anupoju and Marwan Karim"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Using linregpackage}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<style>
body {
  max-width: 1050px;
  margin-left: auto;
  margin-right: auto;
  line-height: 1.6;
}

h1, h2, h3 {
  margin-top: 1.4em;
}

pre {
  overflow-x: auto;
}

img {
  max-width: 100%;
  height: auto;
}
</style>

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.align = "center",
  out.width = "90%"
)
```

# Introduction

The `linregpackage` package implements multiple linear regression using
QR decomposition.

The main function is `linreg()`. It takes a model formula and a data
frame, estimates the regression coefficients, and returns an S3 object
of class `linreg`.

The package also provides methods for inspecting the fitted model,
including `print()`, `coef()`, `resid()`, `pred()`, `summary()`, and
`plot()`.

# Load the package

```{r}
library(linregpackage)
```

# Example data

In this vignette, the built-in `iris` dataset is used.

```{r}
head(iris)
```

We will model `Petal.Length` using `Sepal.Length` and `Sepal.Width` as
explanatory variables.

# Fit a linear regression model

A regression model can be fitted using `linreg()`.

```{r}
model <- linreg(
  Petal.Length ~ Sepal.Length + Sepal.Width,
  data = iris
)
```

The resulting object has class `linreg`.

```{r}
class(model)
```

The function internally creates a design matrix from the supplied
formula and estimates the regression coefficients using QR
decomposition.

# Print the fitted model

The fitted model can be printed using `print()`.

```{r}
print(model)
```

The output shows the original model call and the estimated regression
coefficients.

# Extract regression coefficients

The estimated coefficients can be extracted with `coef()`.

```{r}
coef(model)
```

The output contains the intercept and one coefficient for each
explanatory variable in the model.

# Extract residuals

Residuals represent the difference between the observed response values
and the fitted values from the model.

They can be extracted using `resid()`.

```{r}
head(resid(model))
```

# Extract predicted values

The fitted or predicted values can be returned using `pred()`.

```{r}
head(pred(model))
```

# Model summary

The `summary()` method presents information about the fitted regression
model.

It includes coefficient estimates, standard errors, t-values, p-values,
the residual standard deviation, and the residual degrees of freedom.

```{r}
summary(model)
```

# Diagnostic plots

The `plot()` method creates two diagnostic plots.

The first is a Residuals vs Fitted plot.

The second is a Scale-Location plot.

```{r, fig.width=7, fig.height=5}
plot(model)
```

These plots can be used to inspect the residual behaviour of the fitted
model.

# Complete workflow

A complete analysis with the package can be performed as follows:

```{r}
model <- linreg(
  Petal.Length ~ Sepal.Length + Sepal.Width,
  data = iris
)

print(model)
coef(model)
head(resid(model))
head(pred(model))
summary(model)
```

The `linreg` object stores the quantities calculated during model
fitting, while the S3 methods provide convenient ways to inspect and
display the model results.
