---
title: "Using linregpackage"
author: "Venkata Balaji Anupoju and Marwan Karim"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Using linregpackage}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<style>
body {
  max-width: 900px;
  margin: 0 auto;
  padding: 25px 35px;
  line-height: 1.6;
}

h1, h2 {
  margin-top: 2em;
}

pre {
  border-radius: 6px;
  padding: 12px;
}

img {
  display: block;
  margin: 25px auto 40px auto;
  max-width: 72%;
  height: auto;
}
</style>

```{r, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.align = "center",
  out.width = "72%"
)
```

# Introduction

The `linregpackage` package implements multiple linear regression using
QR decomposition.

The main function is `linreg()`. It takes a model formula and a data
frame, estimates the regression coefficients, and returns an S3 object
of class `linreg`.

The package also provides methods for inspecting the fitted model,
including `print()`, `coef()`, `resid()`, `pred()`, `summary()`, and
`plot()`.

# Load the package

```{r}
library(linregpackage)
```

# Example data

In this vignette, the built-in `iris` dataset is used.

```{r}
head(iris)
```

We will model `Petal.Length` using `Sepal.Length` and `Sepal.Width` as
explanatory variables.

# Fit a linear regression model

A regression model can be fitted using `linreg()`.

```{r}
model <- linreg(
  Petal.Length ~ Sepal.Length + Sepal.Width,
  data = iris
)
```

The resulting object has class `linreg`.

```{r}
class(model)
```

The function internally creates a design matrix from the supplied
formula and estimates the regression coefficients using QR
decomposition.

# Print the fitted model

The fitted model can be printed using `print()`.

```{r}
print(model)
```

The output shows the original model call and the estimated regression
coefficients.

# Extract regression coefficients

The estimated coefficients can be extracted with `coef()`.

```{r}
coef(model)
```

The output contains the intercept and one coefficient for each
explanatory variable in the model.

# Extract residuals

Residuals represent the difference between the observed response values
and the fitted values from the model.

They can be extracted using `resid()`.

```{r}
head(resid(model))
```

# Extract predicted values

The fitted or predicted values can be returned using `pred()`.

```{r}
head(pred(model))
```

# Model summary

The `summary()` method presents information about the fitted regression
model.

It includes coefficient estimates, standard errors, t-values, p-values,
the residual standard deviation, and the residual degrees of freedom.

```{r}
summary(model)
```

# Diagnostic plots

The `plot()` method creates two diagnostic plots.

The first plot shows the residuals against the fitted values.
The second is a Scale-Location plot based on standardized residuals.

```{r diagnostic-plots, fig.width=6, fig.height=4, fig.align="center", out.width="68%"}
plot(model)
```

These plots can be used to inspect the residual behaviour of the fitted
linear regression model.

# Linkoping University graphical theme

The package also includes a ggplot2 theme based on the graphical profile
of Linkoping University. The theme uses the LiU colour palette together
with a clean layout.

```{r liu-theme, fig.width=6, fig.height=4, fig.align="center", out.width="68%"}
theme_data <- data.frame(
  group = c("A", "B", "C"),
  value = c(4, 7, 5)
)

ggplot2::ggplot(
  theme_data,
  ggplot2::aes(
    x = group,
    y = value,
    fill = group
  )
) +
  ggplot2::geom_col(
    width = 0.6
  ) +
  scale_fill_liu() +
  theme_liu() +
  ggplot2::labs(
    title = "Linkoping University",
    subtitle = "Example using the LiU graphical theme",
    x = NULL,
    y = "Value"
  ) +
  ggplot2::theme(
    legend.position = "none"
  )
```

# Complete workflow

A complete analysis with the package can be performed as follows:

```{r}
model <- linreg(
  Petal.Length ~ Sepal.Length + Sepal.Width,
  data = iris
)

print(model)
coef(model)
head(resid(model))
head(pred(model))
summary(model)
```

The `linreg` object stores the quantities calculated during model
fitting, while the S3 methods provide convenient ways to inspect and
display the model results.
